Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to Monoidal Theories

### 2023-10-19 sections/soundness_and_completeness.tex
Theorem about rewriting

\begin{theorem}
    If $f = g$ in $PROP^{+}(\Sigma)$ then $[f] \Rrightarrow_{\mathcal{S}} [g]$ in $\WellTypedMdaEcospans$.
    \begin{proof}
        \update{
            By induction on the number of applied equations to turn $f$ into $g$?
            E.g. $f = g$ in one step essentially means that $f = l$ for some $l$ in~\ref{fig:string-equations} and $g = r$ (or the other way around).
            Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
            }
    \end{proof}
\end{theorem}

@@

    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$.
    $f \Rightarrow{}_{\mathcal{R}^{+}}$ iff $[f] \Rrightarrow{}_{[\mathcal{R}^{+}]} [g]$.
    \begin{proof}
        \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
            But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
        }
    \end{proof}
\end{theorem}

>>>>

\begin{theorem}
    If $f = g$ in $PROP^{+}(\Sigma)$ then $[f] \Rrightarrow_{\mathcal{S}} [g]$ in $\WellTypedMdaEcospans$.
    \begin{proof}
        \update{
            By induction on the number of applied equations to turn $f$ into $g$?
            E.g. $f = g$ in one step essentially means that $f = l$ for some $l$ in~\ref{fig:string-equations} and $g = r$ (or the other way around).
            Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
            }
        \question{Can we use normal form here as well? $f = g$ means there exists $h$ such that $f =_{SMC^{+}} h$ and $h =_{SMC^{+}}$ and $h$ is of the form $\phi_1 + \ldots + \phi_n$?
                  Then we need to show that $[f] \Rrightarrow_{\mathcal{S}} [h]$ and $[h] \Rrightarrow_{\mathcal{S}} [g]$.
        }
    \end{proof}
\end{theorem}

@@

    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$ and $\langle l, r \rangle \in \mathcal{R}$.
    $a \Rightarrow{}^{+}_{\langle l, r \rangle} b$ iff $[a] \Rrightarrow{}^{+}_{\langle [l], [r] \rangle} [b]$.
    \begin{proof}
        $a \Rrightarrow^{+}_{\langle l, r \rangle} b$ means that either
        \begin{itemize}
            \item $a$ and $b$ are decomposable as
                  \[
                    \scalebox{0.75}{
                        \tikzfig{combinatorial_semantics/prop_rewriting}
                    }
                  \]
                  \question{Here we can probably just reuse the proof going to $\ulcorner a \urcorner$ and then the result follows by functoriality of $[\cdot]$ (or $\llangle \cdot \rrangle$)}
            \item or
                 \[
                    \scalebox{0.75}{
                        \tikzfig{combinatorial_semantics/prop_plus_rewriting}
                    }   
                 \]
                 and $a' \Rightarrow_{\langle l,r \rangle} b'$. This gives us the following decomposition in cospans
                 % https://q.uiver.app/#q=WzAsMjcsWzAsNCwibSJdLFsxLDQsIkFfMSJdLFsyLDQsImkgKyBrIl0sWzMsNCwiOyJdLFs0LDAsImkiXSxbNSwwLCJYXzEiXSxbNiwwLCJqIl0sWzUsMiwiXFx2ZG90cyJdLFs1LDQsIkEnIl0sWzQsNCwiaSJdLFs2LDQsImoiXSxbNSwxLCIrIl0sWzUsMywiKyJdLFs1LDUsIisiXSxbNSw2LCJcXHZkb3RzIl0sWzUsNywiKyJdLFs0LDgsImkiXSxbNSw4LCJYX3ciXSxbNiw4LCJqIl0sWzcsNCwiXFxvdGltZXMiXSxbOCw0LCJrIl0sWzksNCwiayJdLFsxMCw0LCJrIl0sWzExLDQsIjsiXSxbMTIsNCwiaiArIGsiXSxbMTMsNCwiQV8yIl0sWzE0LDQsIm4iXSxbMCwxXSxbMiwxXSxbNCw1XSxbNiw1XSxbOSw4XSxbMTAsOF0sWzE4LDE3XSxbMTYsMTddLFsyMCwyMV0sWzIyLDIxXSxbMjYsMjVdLFsyNCwyNV1d
                \[
                    \adjustbox{scale=0.5, center}{
                % \scalebox{0.75}{
                    \begin{tikzcd}
                    &&&& i & {X_1} & j \\
                    &&&&& {+} \\
                    &&&&& \vdots \\
                    &&&&& {+} \\
                    m & {A_1} & {i + k} & {;} & i & {A'} & j & \otimes & k & k & k & {;} & {j + k} & {A_2} & n \\
                    &&&&& {+} \\
                    &&&&& \vdots \\
                    &&&&& {+} \\
                    &&&& i & {X_w} & j
                    \arrow[from=5-1, to=5-2]
                    \arrow[from=5-3, to=5-2]
                    \arrow[from=1-5, to=1-6]
                    \arrow[from=1-7, to=1-6]
                    \arrow[from=5-5, to=5-6]
                    \arrow[from=5-7, to=5-6]
                    \arrow[from=9-7, to=9-6]
                    \arrow[from=9-5, to=9-6]
                    \arrow[from=5-9, to=5-10]
                    \arrow[from=5-11, to=5-10]
                    \arrow[from=5-15, to=5-14]
                    \arrow[from=5-13, to=5-14]
                \end{tikzcd}}
                \]
                and there is a convex match of $L$ in $A'$.
                It is then also down-closed and conflict reflecting as $A'$ is not hierarchical.
                Need to show that arising pushout complement is a boundary complement.
                \question{A problem is that the above terms does not decompose into a form of $A';C$ where $C$ is a pushout complement.
                This happens because we take composition along the external interfaces and this $A'$ is within a box, so it is not a part of an external interface.
                }
                \Aleksei{discuss this. We can simplify PROP+ rewriting to just a box, but still have a problem with context C}
        \end{itemize}
        \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
            But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
        }
    \end{proof}
\end{theorem}
